% certificate.tex — Companion CERTIFICATE OF CORRECTNESS for the Lean 4 formalization
% accompanying "Uniform rationality of p-adic factorization densities".
%
% STANDALONE document. Compile with pdflatex.
%
% HONESTY DISCIPLINE (non-negotiable). This certificate describes what the Lean
% formalization establishes *modulo an explicit list of assumptions*. It never claims
% "formally verified" without that qualification. Cross-checks are labelled EVIDENCE,
% not proof. The measure wall (no p-adic Haar measure in mathlib v4.31.0, hence no real
% instance of the density) is flagged honestly. Conjectural or partial items
% (r-point / pole-control refinements) are NOT presented as proven.
\documentclass[11pt]{article}

\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{xcolor}
\usepackage[colorlinks=true,linkcolor=blue,urlcolor=blue]{hyperref}
\usepackage{enumitem}

\newcommand{\Qp}{\mathbb{Q}_p}
\newcommand{\OK}{\mathcal{O}_K}
\newcommand{\bbQ}{\mathbb{Q}}
\newcommand{\bbF}{\mathbb{F}}
\newcommand{\code}[1]{\texttt{#1}}
\newcommand{\modulo}{\textsl{modulo}}
\newcommand{\Core}{\textsc{Core}}

\title{Certificate of correctness for the Lean~4 formalization\\
\large accompanying \emph{Uniform rationality of $p$-adic factorization densities}}
\author{Companion to the manuscript}
\date{Lean~4 \code{v4.31.0} / \textsf{mathlib} \code{v4.31.0}\\
Ledger re-verified 2026-06-22}

\begin{document}
\maketitle

\begin{abstract}
This is the ``more certainty than usual'' companion to the manuscript and the released
Lean code in \code{lean\_urat/}. We are exact about what the formalization establishes.
The capstone theorems are genuine Lean \code{theorem}s; their machine-checked
\code{\#print axioms} footprints are recorded here verbatim, regenerated by the audit
harness \code{LeanUrat/AxChk\_baseline.lean}. The development is \emph{verified \modulo{}
an explicit list} of assumptions carried as structure fields, theorem hypotheses, and one
named literature-cited axiom: \textbf{(A1)} the minimal box-wise $p$-adic Haar / equidistribution
identification, \textbf{(A2)} the Gu\`ardia--Montes--Nart (GMN) count-shadow hypotheses, and
\textbf{(A3)} the tame functional equation. It is \textbf{not} an unconditional
machine-checked proof: \textsf{mathlib} \code{v4.31.0} has no $p$-adic Haar measure, so
the density itself cannot be constructed in Lean --- the only inhabited instance is a
trivial degree-$0$ witness. We state this limitation --- the \emph{measure wall} ---
plainly. The accompanying symbolic and finite-field computations are presented as
\emph{evidence}, never as part of the formal proof; they are runnable
(Section~\ref{sec:crosschecks}).
\end{abstract}

\tableofcontents

\section{Purpose: what this certificate is, and what it is not}\label{sec:purpose}

\paragraph{What it is.} A faithful record of a \emph{machine-checked-\modulo{}-explicit-axioms}
formalization plus a set of \emph{independent cross-checks}. Concretely:
\begin{itemize}[leftmargin=1.4em]
\item The paper's two main statements --- \emph{uniform rationality} and \emph{palindromy}
of the projective $p$-adic factorization-type density --- are reduced in Lean to an
explicit minimal interface. The conclusions (rationality, palindromy, the tree
decomposition $\rho=\sum_T C_T$) are Lean \code{theorem}s, never axioms. Everything
available in \textsf{mathlib} is used from \textsf{mathlib}; everything cited from the
literature is a named \code{axiom} or a hypothesis field with a doc-comment citation.
\item The whole development is \emph{conditional} on the assumed ledger (A1)--(A3) of
Section~\ref{sec:axioms}, carried as hypotheses/axioms.
\item Independently, a symbolic Okutsu--Montes (OM) density engine reproduces the
Bhargava--Cremona--Fisher--Gajovi\'c (BCFG) root-count distribution and moments
\emph{exactly} through $n=6$, and a $p$-adic factorization oracle confirms the wild
atoms numerically. These are \emph{evidence} for the recursion, not part of the proof.
\end{itemize}

\paragraph{What it is NOT.} It is \textbf{not} an unconditional proof. The honest
one-line summary:
\begin{quote}
\emph{Verified in Lean~4 \modulo{} $\{$(A1) the minimal box-wise $p$-adic Haar
identification (the \textbf{measure wall}); (A2) the GMN count-shadow hypotheses; (A3)
the tame functional equation$\}$, plus Lean's three core classical axioms.}
\end{quote}
Never ``formally verified'' bare. The paper's density $\rho(n,\sigma;q)$ cannot be
constructed in Lean (Section~\ref{sec:wall}); the conjectural $r$-point / pole-control
refinements are \textbf{not} claimed as proven anywhere here.

\paragraph{How to read the ledger.} For each declaration we list its exact
\code{\#print axioms} footprint. We write \Core{} for
$\{\code{propext},\code{Classical.choice},\code{Quot.sound}\}$ --- Lean's core classical
foundations, equivalent to ordinary classical mathematics, present in essentially every
classical theorem; we do not count them as project assumptions. Any \emph{additional}
name is a project axiom and is part of the explicit \modulo{} list for that declaration.
There is \textbf{no} \code{sorryAx} in any of the $55$ footprints audited
(Section~\ref{sec:capstones}).

\section{The assumed ledger (the explicit \modulo{} list)}\label{sec:axioms}

The development is conditional on exactly the following, beyond \Core. None asserts
rationality, uniformity, or palindromy; each is a structural measure/combinatorial fact
absent from \textsf{mathlib} \code{v4.31.0}, recorded in the established
``recorded-via-hypothesis'' style.

\paragraph{(A1) The minimal box-wise measure axiom --- the single irreducible measure
residue.} Carried as \textbf{two structure fields} of \code{MontesAxiom.MontesData} (not
a global \code{axiom}; hypothesis-discharged, so it does \emph{not} surface in
\code{\#print axioms}):
\begin{itemize}[leftmargin=1.4em]
\item \code{boxHaarEquidist}: per-box Haar volume equals the proved closed value,
\[
  \code{boxMeasure}\ c\ q' \;=\; \code{countCellCoeff}\ c\ q'
  \;=\; (q^{\delta})^{d_S-1}\cdot\code{bb1Value}\ c.\code{polygon}\ q' .
\]
That is: the abstract per-box $p$-adic Haar volume \textbf{equals} the separately
\emph{proved} residual-count $\times$ box-volume value.
\item \code{nodeMeasure\_boxSum}: the node limit \emph{exists} and is \emph{box-additive}
--- $\lim_N \code{stratumCount}\,T\,N / q^{nN}$ equals the box-additive assembly
$\sum_{\text{cells }c}\code{boxMeasure}\,c\,q\cdot\prod_{\text{children}}\code{clusterCount}/\code{countPivot}$.
\end{itemize}
The right-hand sides (the \code{bb1Value} box volume, the $(q^{\delta})^{d_S-1}$ residual
count, the box-additivity arithmetic) are all \emph{proved} (P1--P3 below); only the
\emph{identification} of the abstract Haar volume with that proved value is assumed. This
is the genuine \textbf{measure wall} (Section~\ref{sec:wall}).

\paragraph{(A2) The GMN count-shadow hypotheses.} The remaining \code{MontesData} /
\code{CountingModel} structure fields, none of which carries the conclusion:
\code{shapesOf}, \code{partition} (unique-tree disjoint partition
$\code{decidedCount}\ \sigma\ N=\sum_T\code{stratumCount}\ T\ N$); \code{treeSize},
\code{cells}, \code{cells\_descend} (strict cluster-size drop on descent --- the count
shadow of \code{omCells}/\code{descend\_size\_lt}, supplying well-foundedness);
\code{finiteTermination} (GMN Thm 4.18, $\le n$ leaves); and the elementary analytic-tail
bounds \code{undecided\_le\_split}, \code{sepTail\_tendsto},
\code{discriminantTail\_envelope} (bounds, \emph{not} the conclusion $U_N\to 0$, which is
\emph{derived}; see P6). There is \textbf{no} ``$U_N\to0$'' field and \textbf{no} ``$C_T$
is rational'' field --- those are theorems.

\paragraph{(A3) \code{tame\_functionalEquation} (the one declared \code{axiom},
\code{Interface.lean}).} The genuine \S5 H-tame input: an abstract
\code{DensityFoundation}'s density admits a palindromic rational representative at tame
primes (Del Corso--Dvornicich 2000; paper \S5). This is the \textbf{only} non-core
declared axiom on a capstone footprint; it appears \emph{only} on
\code{goal\_theorem\_montes} (and the witness instance), never on any pure-rationality
constant. It asserts nothing about the count-native $\sum_T C_T$ directly --- the link is
the hypothesis \code{hbridge}.

\paragraph{(A4) \code{hbridge} (hypothesis argument, not an axiom).}
$\forall q'>1,\ F.\code{density}\ n\ \sigma\ q' = \sum_T\code{coeff}\ T\ q'$. The honest
measure-wall identification (``abstract foundation density coincides with the
count-native OM tree-sum''); RHS is an arbitrary function, asserting nothing about a
closed form. Introduces no axiom.

\paragraph{The disconnected measure-route axioms (NOT on the capstone path).} The
module \code{LeanUrat.PadicMeasure} declares genuine \code{axiom}s ---
\code{clusterMeasure}, \code{AX\_cellRecursion}, \code{AX\_columnMeasure}, \code{omCells},
\code{descend}, \code{descend\_size\_lt} (and \code{BB3inf} declares its own parallel
\code{BB3inf.descend}, \code{BB3inf.descend\_size\_lt}). These are the \emph{older measure
route}. They are verified \textbf{absent} from \code{goal\_theorem\_montes} and from every
count-native rationality/decomposition constant (Section~\ref{sec:capstones}); they
survive only on the historical \code{goal\_theorem} and the \code{PadicMeasure}/\code{L6M4}
declarations, kept for the original derivation. Their content: \code{AX\_cellRecursion}
(Igusa--Denef per-cell decomposition; Igusa \emph{Local Zeta Functions} \S7.4, Denef
Invent.\ Math.\ 77 (1984)); \code{AX\_columnMeasure} (Igusa box/shell single-coordinate
Haar value, the per-column input --- the closed form $(1-Q^{-1})^V Q^{-A}$ is then
\emph{proved} in \code{L4.cellVolume\_eq}); \code{omCells} (the finite GMN order-$r$ cell
partition; GMN \S2.5 Def 2.21); \code{descend\_size\_lt} (strict cluster-size drop; GMN
Cor 1.20/3.8, Lemma 3.11(3)).

\section{The capstone ledger (\code{\#print axioms}, verbatim)}\label{sec:capstones}

The footprints below were captured with \code{lake env lean LeanUrat/AxChk\_baseline.lean}
against the built library (regenerated for this certificate; $55$ declarations, $0$
\code{sorryAx}). \Core{} abbreviates $\{\code{propext},\code{Classical.choice},
\code{Quot.sound}\}$.

\subsection{The two Goal theorems (the headline)}

\begin{longtable}{@{}p{0.40\textwidth}p{0.54\textwidth}@{}}
\toprule
\textbf{Declaration} & \textbf{Axiom footprint (verbatim)} \\
\midrule
\endhead
\code{Goal.goal\_theorem\_montes} & \code{[propext, Classical.choice,
tame\_functionalEquation, Quot.sound]} \\
\code{Goal.goal\_theorem} (measure route) & \code{[propext, Classical.choice,
tame\_functionalEquation, Quot.sound, PadicMeasure.AX\_cellRecursion,
PadicMeasure.clusterMeasure, PadicMeasure.descend\_size\_lt, PadicMeasure.omCells]} \\
\bottomrule
\end{longtable}

\noindent\textbf{Reading.} The count-native capstone \code{goal\_theorem\_montes}
carries \emph{only} \code{tame\_functionalEquation} on top of \Core: the entire measure
layer (\code{clusterMeasure}, \code{AX\_cellRecursion}, \code{omCells},
\code{descend\_size\_lt}) is \emph{absent}, and there is no \code{sorryAx}. Its rationality
flows through the pure-counting path
(\code{MontesData.countingDensity\_isRational}); its palindromy through (A3) transferred
by the $\bbQ(t)$ identity theorem (\code{L7}). (A1), (A2) and the bridge (A4) are carried
as \emph{hypotheses} (\code{M}, \code{D}, \code{F}, \code{hbridge}), not axioms. The older
\code{goal\_theorem} retains the full measure route and so carries the measure axioms;
both are kept for the historical derivation.

\subsection{Rationality / decomposition (count-native, only-Montes) --- all \Core}

\begin{longtable}{@{}p{0.60\textwidth}p{0.34\textwidth}@{}}
\toprule
\textbf{Declaration} & \textbf{Footprint} \\
\midrule
\endhead
\code{MontesData.coeff\_isRational} & \Core \\
\code{MontesData.countingDensity\_isRational} & \Core \\
\code{MontesData.countingDensity\_eq\_sum\_coeff} & \Core \\
\code{MontesData.nodeMultiplicativity} (\emph{derived}, was a field) & \Core \\
\code{MontesData.undecidedVanishes} ($U_N\to0$, \emph{derived}) & \Core \\
\code{OMInduction.clusterVol\_isRational} (geometric collapse) & \Core \\
\code{RatFn.isRationalFn\_listSum} / \code{\_div} / \code{\_qpow} & \Core \\
\code{CountingModel.density\_le\_decided\_add\_undecided} & \Core \\
\code{CountingModel.decided\_le\_density} & \Core \\
\code{CountingModel.countingDensity\_is\_squeezed\_limit} & \Core \\
\bottomrule
\end{longtable}

\noindent The decomposition $\rho=\sum_T C_T$, the rationality of each $C_T$, the generic
OM-order induction, the squeeze semantics, and the rational-function closure lemmas all
rest on \Core{} only --- \textbf{no} measure axiom, no \code{tame\_functionalEquation},
no \code{sorryAx}. This is the only-Montes footprint.

\subsection{The proved building blocks --- all \Core}

\begin{longtable}{@{}p{0.60\textwidth}p{0.34\textwidth}@{}}
\toprule
\textbf{Declaration} & \textbf{Footprint} \\
\midrule
\endhead
\code{L7.isPalindromic\_of\_agree}, \code{L7.tame\_to\_all\_primes} & \Core \\
\code{L7.ratfunc\_agree\_of\_infinite}, \code{L7.eval\_reflect\_eq} & \Core \\
\code{L3.card\_monicDegree}, \code{L3.irreducible\_separable} & \Core \\
\code{L3Squarefree.card\_squarefreeMonicDegree} & \Core \\
\code{L3Gauss.gauss\_necklace\_count}, \code{L3Gauss.sum\_e\_mul\_card\_eq\_pow} & \Core \\
\code{L4.cellVolume\_eq}, \code{L4.cellVolume\_pIndependent} & \Core \\
\code{L5fix.selfLoop\_geometric} & \Core \\
\code{M1.projPointCount\_mul}, \code{M1.clusterLaw\_baseChange} & \Core \\
\code{R1.treeDistanceWeight\_values} & \emph{no axioms} (\code{decide}) \\
\code{NestedCollapse.nested\_reproduces\_I2} / \code{\_I3} & \Core \\
\code{NestedCollapse.nested\_block\_lt} & \Core \\
\code{NestedCollapse.geometricLimit\_of\_selfLoop} & \Core \\
\code{MontesAxiom.RescaleC1.selfLoopExponent\_eq\_blockExponent\_sub\_one} & \Core \\
\code{Witness.montes\_bundle\_nonempty} & \Core \\
\code{Witness.montes\_full\_instance} & \Core, \code{tame\_functionalEquation} \\
\code{Witness.DensityFoundation\_nonempty} & \Core, \code{PadicMeasure.clusterMeasure} \\
\bottomrule
\end{longtable}

\noindent \code{DensityFoundation\_nonempty} carries \code{clusterMeasure} because its
witness references the old measure route; it is the OLD $\forall F$ goal witness and is
\textbf{off} the only-Montes path. The complete coupled non-vacuity witness for
\code{goal\_theorem\_montes} is \code{montes\_bundle\_nonempty} (\Core) /
\code{montes\_full\_instance} (\Core, \code{tame\_functionalEquation}).

\subsection{The measure layer (\code{PadicMeasure}, \code{L6M4}, \code{BB3inf}) ---
where the wall lives}

\begin{longtable}{@{}p{0.52\textwidth}p{0.42\textwidth}@{}}
\toprule
\textbf{Declaration} & \textbf{Footprint} \\
\midrule
\endhead
\code{PadicMeasure.clusterMeasure\_isRational\_of\_AX} & \Core, \code{AX\_cellRecursion},
\code{clusterMeasure}, \code{descend\_size\_lt}, \code{omCells} \\
\code{PadicMeasure.clusterMeasureModel\_rec} & \Core, \code{descend\_size\_lt},
\code{omCells} \\
\code{PadicMeasure.T\_BB3}, \code{residualCountFn\_isRational} & \Core \\
\code{PadicMeasure.cellCoeff\_isRational}, \code{bb1\_isRational},
\code{omPivot\_isRational}, \code{natCard\_Fr} & \Core \\
\code{PadicMeasure.descend\_children\_of\_factors} & \Core, \code{omCells} \\
\code{PadicMeasure.cell\_stratum\_of\_residual} & \Core, \code{omCells} \\
\code{L6M4.density\_rational}, \code{L6M4.sum\_clusterVol\_rational} & \Core,
\code{AX\_cellRecursion}, \code{clusterMeasure}, \code{descend\_size\_lt}, \code{omCells} \\
\code{BB3inf.clusterVolume\_rational\_of\_OM} & \Core, \code{BB3inf.descend},
\code{BB3inf.descend\_size\_lt} \\
\code{BB3inf.T\_BB3}, \code{T\_BB1}, \code{bb3CellCount\_isRational} & \Core \\
\code{Interface.clusterVolume\_rational} & \Core, \code{AX\_cellRecursion},
\code{clusterMeasure}, \code{descend\_size\_lt}, \code{omCells} \\
\bottomrule
\end{longtable}

\section{Which parts of the paper's argument are formalized}\label{sec:map}

The table maps the paper's proof skeleton to Lean. \emph{Proved (\Core)} means a Lean
\code{theorem} resting only on \Core. \emph{Assumed} means carried as an (A1)/(A2)/(A3)
hypothesis or measure-route axiom.

\begin{longtable}{@{}p{0.30\textwidth}p{0.42\textwidth}p{0.22\textwidth}@{}}
\toprule
\textbf{Paper (\S/label)} & \textbf{Lean theorem(s)} & \textbf{Status} \\
\midrule
\endhead
Main theorem (\code{thm:main}) & \code{Goal.goal\_theorem\_montes} &
Proved \modulo{} (A1)(A2)(A3) \\
Tree decomposition $\rho=\sum_T C_T$ (\code{eq:tree-decomp}) &
\code{MontesData.countingDensity\_eq\_sum\_coeff} (genuine limit-interchange,
\emph{not} \code{rfl}) & Proved (\Core) \\
BB1 Newton-cell box volume $(1-Q^{-1})^V Q^{-A}$ (\code{lem:bb1}) &
\code{L4.cellVolume\_eq}, \code{MontesAxiom.boxVolume\_eq} & Proved (\Core) \\
BB3 residual count over $\bbF_q$, $=Q^{d_S-1}$ (\code{lem:bb3}) &
\code{MontesAxiom.residualBoxCount} ($d_S\ge2$), \code{PadicMeasure.T\_BB3} & Proved (\Core) \\
Per-shape rationality by OM-order induction (\code{prop:per-shape}) &
\code{OMInduction.clusterVol\_isRational} (well-founded on \code{treeSize}) &
Proved (\Core) \\
Geometric self-loop collapse (the wild absorption, \code{rem:wild-absorbed}) &
\code{NestedCollapse.geometricLimit\_of\_selfLoop}, \code{L5fix.selfLoop\_geometric},
\code{R1.treeDistanceWeight\_values} & Proved (\Core) \\
Box additivity / per-node multiplicativity &
\code{MontesAxiom.clusterCount\_boxSum}, \code{nodeMultiplicativity} (derived) &
Proved (\Core) \\
$U_N\to0$ (undecided pool vanishes) &
\code{MontesData.undecidedVanishes} (squeeze) & Proved (\Core) \\
Monic$\to$projective bridge (\code{sec:bridge}) &
\code{M1.projPointCount\_mul}, \code{clusterLaw\_baseChange} & Proved (\Core) \\
Uniform rationality corollary (\code{cor:rationality}) &
\code{MontesData.countingDensity\_isRational}, \code{coeff\_isRational} & Proved (\Core) \\
Palindromy at wild primes (\code{prop:palindromy}) &
\code{L7.tame\_to\_all\_primes}, \code{isPalindromic\_of\_agree} (via $\bbQ(t)$ identity
theorem) & Proved (\Core), uses (A3) \\
\midrule
\multicolumn{3}{@{}l}{\emph{The measure inputs (assumed, not formalized as proofs):}} \\
$p$-adic integration / Haar normalization (\code{sec:padic-integration}) &
(A1) \code{boxHaarEquidist} + \code{nodeMeasure\_boxSum}; measure route
\code{AX\_cellRecursion}, \code{AX\_columnMeasure} & \textbf{Assumed} \\
GMN OM-tree machinery (\code{hyp:gmn}, \code{sec:om}) &
(A2) \code{cells}, \code{cells\_descend}, \code{finiteTermination}; measure route
\code{omCells}, \code{descend\_size\_lt} & \textbf{Assumed} \\
\bottomrule
\end{longtable}

\noindent The proof \emph{architecture} --- the tree decomposition, BB1, BB3, the
OM-order induction with geometric collapse, the $U_N\to0$ squeeze, the monic$\to$projective
bridge, and the palindromy transfer --- is formalized as \Core{} Lean theorems. The
\emph{measure inputs} (existence and normalization of the $p$-adic Haar volume; the GMN
cell/descent structure) are assumed.

\section{The measure wall (the honest limitation)}\label{sec:wall}

\textsf{mathlib} \code{v4.31.0} provides no $p$-adic Haar measure (no
\code{MeasureSpace}/\code{haarMeasure} on \code{PadicInt} or any DVR; a search of
\code{Mathlib/MeasureTheory} for \code{Padic} is empty). Consequently:
\begin{itemize}[leftmargin=1.4em]
\item The paper's actual density $\rho(n,\sigma;q)$ --- a genuine $p$-adic Haar volume ---
\textbf{cannot be constructed} in Lean. The capstones quantify over an abstract
\code{DensityFoundation}/\code{CountingModel}; they are \emph{conditional} theorems.
\item Non-vacuity \emph{is} witnessed: the full coupled \code{goal\_theorem\_montes}
bundle has a sorry-free inhabitant (\code{Witness.montes\_bundle\_nonempty} /
\code{montes\_full\_instance}), with non-trivial density $g_\sigma\equiv 1$ and (A1)
holding by \code{rfl} (\code{boxMeasure}$:=$\code{countCellCoeff}). \emph{But} the only
inhabited instance is a trivial degree-$0$ witness ($q{=}2$, $n{=}0$:
\code{trivM}/\code{trivD}/\code{trivF}), \textbf{not} the paper's $\rho$. Identifying any
constructed instance with $\rho$ is exactly the un-crossed wall.
\item Concretely, the wall is (A1): that the per-box $p$-adic Haar volume equals the
separately \emph{proved} lattice/finite-field value. The right-hand sides of these
identities are proved; only their \emph{identification with the Haar volume} is assumed.
\end{itemize}
We therefore describe the formalization as \emph{formally verified \modulo{} the explicit
list of Section~\ref{sec:axioms}}, never as ``formally verified'' bare. The
$r$-point/pole-control refinements and any conjectural higher-order item are
\textbf{not} claimed as proven here.

\section{Pre-existing flagged \code{sorry}s (off every capstone footprint)}\label{sec:sorry}

Two declarations in \code{LeanUrat.L3} carry a flagged \code{sorry}:
\code{L3.card\_squarefreeMonicDegree} and \code{L3.gauss\_necklace\_count}. Both require
generating-function / M\"obius identities absent from \textsf{mathlib} \code{v4.31.0}.
They are \textbf{re-proved sorry-free} as standalone theorems:
\code{L3Squarefree.card\_squarefreeMonicDegree} (the $A\cdot B^2$ unique-factorization
count) and \code{L3Gauss.gauss\_necklace\_count} (the necklace identity plus M\"obius
inversion). Because \code{L3Squarefree} and \code{L3Gauss} both \emph{import} \code{L3},
wiring the proved statements back into \code{L3} would create an import cycle; the cycle
blocks the wiring, so the two \code{L3} \code{sorry}s remain. This is benign: the
count-native capstone path consumes the \emph{sorry-free} versions (via \code{T\_BB3} /
\code{residualBoxCount}), never the \code{L3} stubs. No capstone \code{\#print axioms}
footprint contains \code{sorryAx} (verified: $0$ occurrences across all $55$ audited
declarations).

\section{Independent cross-checks (EVIDENCE, not proof)}\label{sec:crosschecks}

The computations below were run \emph{outside} Lean and are \emph{evidence} that the
closed-form arithmetic the Lean modules prove symbolically matches independent
computation; they are not part of the formal proof. Two reproducible scripts (system
\code{python3} + \code{sympy} \code{1.13.3}):
\begin{itemize}[leftmargin=1.4em]
\item \code{uniform-rationality/\_cert\_crosschecks.py} --- $25$ exact checks
(sympy identities + exhaustive $\mathrm{GF}(Q)$ enumeration; no floats). All pass.
\item \code{uniform-rationality/\_reconcile\_om\_bcfg.py} --- the symbolic OM density
engine vs BCFG, $n=2{\dots}6$; output \code{results/reconcile\_om\_bcfg.json}
(\code{all\_match: true}).
\end{itemize}

\subsection{Closed-form arithmetic the Lean modules prove (exact)}

\begin{longtable}{@{}p{0.40\textwidth}p{0.42\textwidth}p{0.10\textwidth}@{}}
\toprule
\textbf{Lean claim} & \textbf{Independent check} & \textbf{Result} \\
\midrule
\endhead
\code{R1.treeDistanceWeight}, \code{L5fix.selfLoopExponent}: $w(s)=s(s{+}1)/2-1$ &
$w(2,3,4,5)=2,5,9,14$ (exact) & match \\
C1 \code{selfLoopExponent\_eq\_blockExponent\_sub\_one}:
$w(s)=\big(s+\binom{s}{2}\big)-1$ & $s=0,\dots,7$ (exact $\mathbb{N}$) & match \\
\code{NestedCollapse.closedI2}: $I_2=q/(q+1)$ via $a/(1-r)$,
$a=(q-1)/q$, $r=q^{-2}$ & \code{sympy.simplify} $\equiv 0$ & match \\
\code{NestedCollapse.closedI3}:
$I_3=\dfrac{q^3(q^2-q+1)}{(q+1)\Phi_5}$ from \code{nestedRHS} &
\code{sympy.simplify} $\equiv 0$ & match \\
\code{L4.cellVolume\_eq}: column product $=(1-Q^{-1})^VQ^{-A}$ &
$(V,A)\in\{(1,1),(2,3),(3,5)\}$, exact & match \\
\code{L3Squarefree.card\_squarefreeMonicDegree} ($d_S\ge2$):
$S(m)=Q^m-Q^{m-1}$ & $\mathrm{GF}(Q)$ enum, $Q\in\{2,3\}$, $d=2,3$ & match \\
\code{L3Squarefree.sum\_identity} (recursion $(\ast)$):
$Q^n=\sum_{2j\le n}S(n{-}2j)Q^j$ & $Q\in\{2,3,5\}$, $n=0,\dots,8$ exact & match \\
\code{L3Gauss}: $d\,N_d=\sum_{e\mid d}\mu(e)Q^{d/e}$,
$\sum_{e\mid d}eN_e=Q^d$ & $\mathrm{GF}(Q)$ enum, $Q\in\{2,3,5\}$, $d\le4$ & match \\
\code{M1.projPointCount\_mul}: $(q-1)\sum_{i=0}^n q^i=q^{n+1}-1$ &
$n=0,\dots,5$ exact & match \\
\code{T\_BB3}/\code{residualBoxCount} ($d_S\ge2$): $\#\{$non-squarefree
monic $\deg d\}=Q^{d-1}$ & $\mathrm{GF}(Q)$ enum, $Q\in\{2,3\}$, $d=2,3$ & match \\
\bottomrule
\end{longtable}

\noindent\textbf{Honest note on hypotheses.} The closed forms $S(m)=Q^m-Q^{m-1}$ and
$\#\{$non-squarefree$\}=Q^{d-1}$ are the \emph{Lean theorem statements}, which carry the
hypothesis $2\le m$ (resp.\ $2\le d_S$): at $m=1$ all $Q$ degree-$1$ monics are
squarefree, so $S(1)=Q\ne Q-1$ and the non-squarefree count is $0\ne Q^0$. The cross-check
mirrors that hypothesis exactly (it tests $d\ge2$). The recursion $(\ast)$ holds for all
$n$ (using the correct $S(0)=1$, $S(1)=Q$, $S(m{\ge}2)=Q^m-Q^{m-1}$). All
$\mathrm{GF}(Q)$ rows are genuine exhaustive enumerations of monic polynomials over a prime
field, so they verify the finite-field counts directly. No floating-point step.

\subsection{The symbolic OM engine reproduces BCFG root-count moments through $n=6$}

The OM density engine (\code{uniform-rationality/om\_density\_engine.py}) assembles the
\emph{full per-type} symbolic distribution $\rho(n,\sigma;q)$ (e.g.\ all $34$ types at
$n=6$, including the first deep-wild OM-order-$3$ branch), aggregates to the root-count
distribution $P(n,r;q)$, and compares to BCFG's independently-derived $P(n,r;q)$
(arXiv:2101.09590 Thm 1; generating-function inversion, methodologically independent of
the OM/cluster engine). Every comparison is an \emph{exact} rational-function identity
\code{sympy.simplify(OM$-$BCFG)$=0$}, re-run for this certificate
(\code{results/reconcile\_om\_bcfg.json}, \code{all\_match: true}).

\begin{longtable}{@{}cccc@{}}
\toprule
$n$ & \#types & support $r$ & all $P(n,r;q)$ match BCFG \\
\midrule
\endhead
$2$ & $3$  & $\{0,2\}$       & yes ($3$ cells) \\
$3$ & $5$  & $\{0,1,3\}$     & yes ($4$ cells) \\
$4$ & $11$ & $\{0,1,2,4\}$   & yes ($5$ cells) \\
$5$ & $17$ & $\{0,1,2,3,5\}$ & yes ($6$ cells) \\
$6$ & $34$ & $\{0,1,2,3,4,6\}$ & yes ($7$ cells) \\
\bottomrule
\end{longtable}

\noindent Additional exact ($\code{simplify(diff)}=0$) checks in the same run:
$\sum_r P(n,r;q)=1$ at every $n$; $\sum_\sigma \rho(n,\sigma;q)=1$ (the exhaustive
enumeration gate); $E[N]=1$ for all $n$; the stabilized moments
$E[N^2]=(2q^4+q^3+3q^2+q+2)/\Phi$ and $E[(N)_2]=(q^2+1)^2/\Phi$ for $n\ge3$; and the
factorial moments $E[(N)_d]$ matching BCFG's $d$-th-moment stabilization
($n\ge2d-1$). The moments are \emph{also} derivable via the Vandermonde/Selberg
local-integral route (Track A), matching BCFG through $d\le4$
(\code{trackA\_highd\_moments.py}).

\paragraph{Per-type oracle (closes the within-$r$ caveat).} $P(n,r;q)$ is a per-$r$
aggregate; two engines could agree on every $P(n,r)$ while disagreeing on individual
type-densities. A PARI \code{factorpadic} projective oracle probed the \emph{individual}
densities $\rho(6,\sigma;p)$ at $p\in\{2,3,5,7\}$, $N\sim30$k/prime: $0$
observed-but-unpredicted types, overall $\max|z|=3.66$ (sign-flipping across primes ---
MC Poisson noise, not a systematic error), and every deep-wild order-$3$ frontier type
matches within noise. This per-type agreement is empirical (MC), not a proof.

\subsection{$n\le3$ exact closed forms vs BCFG, and the wild atoms}

\begin{longtable}{@{}p{0.42\textwidth}p{0.40\textwidth}p{0.10\textwidth}@{}}
\toprule
\textbf{Claim (paper / engine)} & \textbf{Value, vs BCFG} & \textbf{Result} \\
\midrule
\endhead
$\rho(3,\{(1,1)^3\};q)$ (all rational roots) & $\dfrac{(q^2+1)^2}{6\,\Phi}$ &
match \\
$P(3,0;q)$ (no rational root) & $\dfrac{(q^2+1)^2}{3\,\Phi}$ & match \\
$P(3,1;q)$ & $\dfrac{q^4+2q^3+2q+1}{2\,\Phi}$ & match \\
$n=3$ five type densities (paper Ex.) & all palindromic, $\sum=1$ & match \\
First wild atom $n=4$, slope-$\tfrac12$, conditional law &
$P((4,1)\mid\text{insep.})=\dfrac{q}{q+1}$ & match \\
\quad vs recorded oracle (incl.\ \textbf{wild $p=2$ Artin--Schreier}) &
obs $0.6664$ at $p{=}2$ vs $q/(q{+}1)=0.6667$ ($N{=}80$k) & match \\
\quad oracle at $p=3,5,7,11,13$ & $0.7515,0.8329,0.8756,0.9147,0.9289$ &
within $4\sigma$ \\
Deep-wild $n=6$ order-$3$ cluster law $\beta_{(6,1)}(q)$ &
$\dfrac{q^2(q-1)(q^2+1)}{q^5-1}$,\ $\beta_{(6,1)}(2)=\tfrac{20}{31}$ & match \\
\quad rejects naive guess $q/(q{+}1)=\tfrac23$ & oracle $\approx 11\sigma$ rejection
& reject \\
\bottomrule
\end{longtable}

\noindent The $n=4$ wild $p=2$ row is the deepest evidence: the char-dependent geometry
(a punctured parabola $\bar b^2=4\bar a$ at odd $p$ versus a punctured line $\bar b=0$,
$R=(y+\sqrt{\bar a})^2$, inseparable at $p=2$) yields the \emph{same} conditional law
$q/(q+1)$. The recorded oracle ($\code{results/n4wild\_insep\_p\{2,3,5,7,11,13\}\_s30*.json}$,
$N=80$k/prime aggregated) matches the exact prediction within MC noise at every prime,
including wild $p=2$ at deviation $0.0003$ ($4\sigma$ band $0.0067$). The
PARI \code{factorpadic} binary is unavailable in this build environment; these rows reuse
the previously recorded oracle data as the empirical anchor and re-verify the symbolic
predictions and the comparison, which are reproducible.

\section{Release module list}\label{sec:modules}

The library \code{LeanUrat} (target of \code{lake build LeanUrat}) comprises:
\code{Interface}, \code{PadicMeasure}, \code{OMInduction}, \code{RatFn},
\code{CountingModel}, \code{MontesAxiom}, \code{Decomposition}, \code{NestedCollapse},
\code{L3}, \code{L3Squarefree}, \code{L3Gauss}, \code{L4}, \code{L5fix}, \code{L6M4},
\code{L7}, \code{M1}, \code{R1}, \code{BB3inf}, \code{Goal}, \code{Witness}, \code{Smoke}
(+ \code{Basic}, the lake template placeholder; \code{AxChk\_baseline}, the audit harness
that regenerates Section~\ref{sec:capstones}).

\section*{Provenance statement}
The axiom footprints in Section~\ref{sec:capstones} are reproduced verbatim from
\code{\#print axioms} run on the built library via
\code{lake env lean LeanUrat/AxChk\_baseline.lean} ($55$ declarations, $0$ \code{sorryAx}).
The cross-checks in Section~\ref{sec:crosschecks} are exact symbolic / exhaustive
finite-field computations (\code{\_cert\_crosschecks.py}, $25/25$ pass) and the OM--BCFG
reconciliation (\code{\_reconcile\_om\_bcfg.py}, \code{all\_match: true}), run
independently of Lean. We have stated the measure wall and the two flagged \code{L3}
\code{sorry}s plainly; no conjectural item ($r$-point / pole control) is presented as
proven.

\end{document}
